### 2017-06-28 motivation.tex
Fig. 4 shrunk to fit 12 pages.
Format changed to PACMPL.

In this section, we motivate our ideas via an example written in
\txnimp - a C-like imperative language equipped with a \C{txn} lexical
block defining a transaction scope. Each \C{txn} block is associated
with an identifier in angle braces that uniquely identifies the
transaction. We use Hoare triple notation to annotate programs with
pre- and post- conditions.

\begin{figure}
\centering
$\{\texttt{B}=\texttt{k} \conj \texttt{k}\ge\texttt{a1+a2}\}$
\begin{tabular}{l||l}
\begin{txnimpcode}
  txn$\langle$'Wd1'$\rangle${
    if (B $\ge$ a1) {
      B := B - a1
    }
  }
\end{txnimpcode}
&
\begin{txnimpcode}
  txn$\langle$'Wd2'$\rangle${
    if (B $\ge$ a2) {
      B := B - a2
    }
  }
\end{txnimpcode}
\\
\end{tabular}
$\{\texttt{B}=\texttt{k-a1-a2}\}$
\caption{\small Concurrent withdraw transactions}
\label{fig:motiv-eg-1}
\vspace*{-10pt}
\end{figure}

@@

\begin{figure}
\centering
\subcaptionbox {
  \label{fig:hb-eg-1}
} [
  0.25\columnwidth
] {
  \includegraphics[scale=0.5]{Figures/hb-eg-1}
}
\subcaptionbox {
  \label{fig:hb-eg-2}
} [
  0.25\columnwidth
] {
  \includegraphics[scale=0.5]{Figures/hb-eg-2}
}
\subcaptionbox {
  \label{fig:hb-eg-3}
} [
  0.25\columnwidth
] {
  \includegraphics[scale=0.5]{Figures/hb-eg-3}
}
\subcaptionbox {
  \label{fig:hb-eg-4}
}{
  \includegraphics[scale=0.5]{Figures/hb-eg-4}
}

>>>>

In this section, we motivate our ideas via an example written in
\txnimp - a functional language equipped with a \C{DB} monad to
compose database computations. Besides the usual \C{bind} and
\C{return} combinators, monad offers a \C{txn\_do} combinator that
executes a database computation and returns the result.

\begin{figure}
\centering
\begin{tabular}{l||l}
\begin{ocaml}
let new_order_txn d_id c_d_id c_id
      (ireqs: item_req list) = atomically do
  dist <- SQL.select1 District
               (fun d -> d.d_id = d_id);
  let o_id = dist.d_next_o_id;
  SQL.update District 
    (fun d -> 
      {d with d_next_o_id =d_next_o_id + 1})
    (fun d -> d.d_id = d_id );
  SQL.insert Order {o_id=o_id;  o_d_id=d_id; 
                    o_c_id=c_id; o_ol_cnt=S.size ireqs; };
  SQL.insert New_order {no_o_id=o_id; no_d_id=d_id};
  foreach ireqs @@ fun ireq -> do
    stk <- SQL.select1 Stock
             (fun s -> s.s_i_id = ireq.ol_i_id);
    let s_qty' = if stk.s_qty >= ireq.ol_qty + 10 
                then stk.s_qty - ireq.ol_qty
                else stk.s_qty - ireq.ol_qty + 91;
    SQL.update Stock
       (fun s -> {s with s_qty = s_qty'})
       (fun s -> s.s_i_id = ireq.ol_i_id);
    SQL.insert Order_line {ol_o_id=o_id; ol_d_id=d_id; 
        ol_i_id=ireq.ol_i_id; ol_qty=ireq.ol_qty}

\end{ocaml}
&
\begin{ocaml}
  txn$\langle$'Wd2'$\rangle${
    if (B $\ge$ a2) {
      B := B - a2
    }
  }
\end{ocaml}
\\
\end{tabular}
\caption{\small Concurrent withdraw transactions}
\label{fig:motiv-eg-1}
\vspace*{-10pt}
\end{figure}

@@

\begin{figure}
\centering
\subcaptionbox {
  \label{fig:hb-eg-1}
} [
  0.25\columnwidth
] {
  \includegraphics[scale=0.45]{Figures/hb-eg-1}
}
\subcaptionbox {
  \label{fig:hb-eg-2}
} [
  0.25\columnwidth
] {
  \includegraphics[scale=0.45]{Figures/hb-eg-2}
}
\subcaptionbox {
  \label{fig:hb-eg-3}
} [
  0.25\columnwidth
] {
  \includegraphics[scale=0.45]{Figures/hb-eg-3}
}
\subcaptionbox {
  \label{fig:hb-eg-4}
}{
  \includegraphics[scale=0.45]{Figures/hb-eg-4}
}
